The SMT Competition 2015--2018

### 2018-11-23 2018/report/SMT-COMP.tex
Minor

The SMT Library classifies benchmarks according to their \emph{logic},
that is, according to the specific combination of background theories
that the benchmark uses.  Logic names are composed of letter groups
that refer to these background theories: {A} or {AX} for the theory of
arrays, {BV} for the theory of bit-vectors, {DT} for algebraic
datatypes, {FP} for the floating-point theory, {IA} for integer
arithmetic, {RA} for real arithmetic, {IRA} for mixed integer and real
arithmetic, {S} for the theory of strings, {UF} for uninterpreted
functions (i.e., free symbols).  Additionally, logics may impose
further syntactic restrictions, such as the absence of quantifiers
({QF\_}), the restriction to difference logic over the
integers~({IDL}) or reals ({RDL}), or to the (non-)linear fragment of
arithmetic ({L} or {N}, respectively, before {IA}, {RA} or {IRA}).
For instance, the logic QF\_UFLIA contains quantifier-free formulas
with uninterpreted functions over linear integer arithmetic.  These
restrictions are typically motivated by the existence of efficient
decision procedures for certain syntactic fragments.

@@

There are slightly fewer divisions than logics in the SMT Library
(Table~\ref{table:smtlib}) because not every logic contains benchmarks
eligible for the competition.  Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, 11,966 benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division) whose precise semantics were
being debated at the time.  Additionally in 2016, one benchmark was
excluded because it contained incorrect status information.  In 2017,
1,104 benchmarks were excluded because their logic had been classified
incorrectly, and 8 benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the unsat-core
track because of lacking tool support.  In 2018, 3,603 benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and 5
benchmarks were excluded because of missing status information.

@@

Also shown in Figure~\ref{fig:job-pairs} is the total wall-clock time
(summed over all job pairs) for the competition.  This has more than
doubled, from 3.5~years in~2015 to 7.6~years in~2018.  (These are
solver run-times only.  Time that the cluster spent on pre- and
post-processing of benchmarks and solver output or on other auxiliary
tasks is not included.)  The increase in~2016 is largely due to the
inclusion of non-incremental benchmarks with unknown status; these
were evaluated with a wall-clock time limit of 10~minutes.  In~2017,
when these benchmarks had become eligible for the main track,
main-track job pairs were first run with a time limit of 10~minutes;
when it became apparent to the organizers that there were sufficient
resources to increase this limit, job pairs that had timed out were
re-run with a time limit of 20~minutes.  (In this way, about 2.9 years
of wall-clock time in 2017 were spent on job pairs that were then
re-run.  This time is not included in Figure~\ref{fig:job-pairs}.)
Table~\ref{table:time-limits} shows the wall-clock time limits that
were imposed on job pairs for each track and year.

\begin{table}
  \caption{Wall-clock time limits.}
  \label{table:time-limits}
  \centering
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    & \multicolumn{1}{c|}{2015} & \multicolumn{1}{c|}{2016} & \multicolumn{1}{c|}{2017} & \multicolumn{1}{c|}{2018} \\
    \hline
    Main track        & 40~min & 40~min & 20~min\protect\footnotemark & 20~min \\
    Application track & 40~min & 40~min & 40~min\phantom{$^{1.}$} & 40~min \\
    Unsat-core track  &     -- & 40~min & 40~min\phantom{$^{1.}$} & 40~min \\
    \hline
  \end{tabular}
\end{table}
\footnotetext{These job pairs were first run with a time limit of
  10~minutes.  See the text for details.}

@@

Table~\ref{table:participants} lists the solvers that participated
between~2015 and~2018, together with their submitter(s) and
institutional affiliations (which may have changed over time).  Note
that most solvers were developed by a group of people; the table does
not list all contributors but only submitters.
%
Participants are also encouraged to submit a short system description
of their solver, including a list of its authors and an explanation of
the SMT solving approach used.  System descriptions are published on
the SMT-COMP website.  They were submitted for about 60\,\% of all
solver versions since 2015.

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{11.7cm}|} %TODO: fix width
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC~\cite{DBLP:conf/cav/BraytonM10} & Valeriy Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE~\cite{DBLP:conf/cade/GieslBEFFOPSSST14} & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector~\cite{DBLP:conf/cav/NiemetzPWB18} & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI~\cite{colibri} & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo~\cite{ctrl-ergo} & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3~\cite{cvc3}  & Kshitij Bansal
                        (New York University)\\
    CVC4~\cite{DBLP:conf/cav/BarrettCDHJKRT11}  & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP~\cite{maplestp} & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink~\cite{minkeyrink} & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2~\cite{opensmt2} & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB~\cite{prob}  & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B~\cite{q3b}    & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT~\cite{rasat} & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog~\cite{redlog} & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT~\cite{smt-rat} & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol~\cite{} & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT~\cite{} & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP~\cite{}        & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt~\cite{}     & Masahiro Sakai
                        (Toshiba Corporation)\\
    Vampire~\cite{}    & Giles Reger
                        (University of Manchester)\\
    veriT~\cite{DBLP:conf/cade/BoutonODF09} & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat~\cite{}       & Martin Velez
                        (University of California, Davis)\\
    Yices2~\cite{}     & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3~\cite{DBLP:conf/tacas/MouraB08} & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}
\end{table}

@@

The competition-wide ranking for the main track was retained and
refined after~2014.  Table~\ref{table:competition-wide-ranking} shows
the best three solvers according to this ranking for 2015--2018.  Also
shown is Z3, which ranked highly each year but was running hors
concours.  Sequential and parallel performance are reported
separately.
%
Note that scores are not directly comparable between different years:
the set of benchmarks changes from one year to the next, and also the
formula to compute the score changed (see below).  Moreover, readers
should keep in mind that this ranking---like any method that maps the
performance of solvers for a broad range of logics to a
one-dimensional value---is inherently limited.  We refer to the
individual division rankings (Section~\ref{sec:division-rankings}) for
more detailed information.

\begin{table}
  \caption{Competition-wide ranking for the main track: best solvers (by year).}
  \label{table:competition-wide-ranking}
  \centering
  \scalebox{.89}{\begin{tabular}{|c|p{2.2cm}|r|r|p{0cm}|c|p{2.2cm}|r|r|}
    \cline{1-4} \cline{6-9}
    \multicolumn{4}{|c|}{2015} & & \multicolumn{4}{c|}{2016} \\ \cline{1-4} \cline{6-9}
    Rank & Solver & Score  & Score  & & Rank & Solver & Score & Score \\
         &        & (sequential) & (parallel) & & & & (sequential) & (parallel) \\ \cline{1-4} \cline{6-9}
         & [Z3]              & 159.36 & 159.36 & &    & [Z3]   & 185.09 & 185.09 \\
      1. & CVC4              & 144.67 & 144.74 & & 1. & CVC4   & 180.95 & 181.19 \\
      2. & CVC4(exp)\footnotemark & 140.47 & 140.51 & & 2. & Yices2 & 119.29 & 119.29 \\
      3. & Yices             & 101.91 & 101.91 & & 3. & veriT  &  75.11 &  75.11 \\ \cline{1-4} \cline{6-9}
  \end{tabular}}

@@

After 13 years, the SMT Competition is a well-established event that
garners wide interest from developers and users of SMT solvers.  The
competition continues to contribute to its original goals, including
adoption of the SMT-LIB format and sparking further advances in SMT,
and provides a yardstick of performance for SMT solvers.

The last FLoC Olympiad was a period of significant growth for the SMT
Competition.  The average number of solver submissions increased
from~12 (for 2005--2014) to~26 (for 2015--2018).  The number of
SMT-LIB benchmarks used in the competition increased from~77,352
in~2014~\cite{CDW14} to 342,498 in 2018.  This is partly because new
benchmarks were added to the SMT Library, but also because benchmarks
with unknown status are now eligible for the competition, and because
the earlier culling of benchmarks was abandoned.  Since the
re-introduction of the unsat-core track in~2016, the competition again
has three separate tracks.  This, together with new logics in the SMT
Library, has brought the total number of competition divisions to 115
in~2018, up from just 40 divisions in~2014~\cite{CDW14}.  The number
of job pairs increased from 339,714 in~2014 to~1,776,062~in 2018, and
the total wall-clock time in~2018 exceeded 7.6~years.  By these
measures, the SMT Competition is one of the largest competitions in
automated reasoning.

@@

Despite the substantial amount of computational resources available to
the competition through StarExec, the competition has reached a size
where future organizers will have to carefully consider the number of
participants and benchmarks each year, to determine whether the
current time limits remain tenable.  (Otherwise, reducing the time
limits would probably be a better solution than reverting to the
pre-2014 practice of benchmark culling.)

Further development of the competition should be undertaken with care,
keeping in mind that major changes require not just the competition
organizers, but also participants to adapt.  Immediate action items
for next year's organizers are
\begin{itemize}
\item the inclusion of {\tt check-sat} commands with unknown status in
  the application track; and
\item the definition of competition-wide rankings for the application
  and unsat-core track.
\end{itemize}

@@

Organizing the competition is a non-trivial amount of work.  Much of
this happens behind the scenes.  For instance, the competition tools
(such as the benchmark scrambler and trace executor) often require
updating when new divisions are added, or when changes are made to
StarExec.  One could imagine adding further tracks (e.g., for
proof-producing SMT solvers, or for variants such as max-SMT) to the
competition, but the effort required to support such tracks will need
to be balanced against their added value.

To ensure that the competition results are practically relevant, it is
important that the SMT Library contains benchmarks that cover a wide
range of applications.  To this end, we again encourage users of SMT
solvers to submit new and interesting benchmarks to the library,
especially for underrepresented logics.  This is not just a valuable
contribution to the community, but also---since these benchmarks will
then be used in the competition---a straightforward way to get solver
developers to tune their solvers accordingly.

>>>>

The SMT Library classifies benchmarks according to their \emph{logic},
that is, according to the specific combination of background theories
that the benchmark uses.  Logic names are composed of letter groups
that refer to these background theories: {A} or {AX} for the theory of
arrays, {BV} for the theory of bit-vectors, {DT} for algebraic
datatypes, {FP} for the floating-point theory, {IA} for integer
arithmetic, {RA} for real arithmetic, {IRA} for mixed integer and real
arithmetic, {S} for the theory of strings, {UF} for uninterpreted
functions (i.e., free symbols).  Additionally, logics may impose
further syntactic restrictions, such as the absence of quantifiers
({QF\_}), the restriction to difference logic over the
integers~({IDL}) or reals ({RDL}), or to the (non-)linear fragment of
arithmetic ({N} or {L}, respectively, before {IA}, {RA} or {IRA}).
For instance, the logic QF\_UFLIA contains quantifier-free formulas
with uninterpreted functions over linear integer arithmetic.  These
restrictions are typically motivated by the existence of efficient
decision procedures for certain syntactic fragments.

@@

There are slightly fewer divisions than logics in the SMT Library
(Table~\ref{table:smtlib}) because not every logic contains benchmarks
eligible for the competition.  Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, 11,966 benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division) whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
1,104 benchmarks were excluded because their logic had been classified
incorrectly, and 8 benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the unsat-core
track because of lacking tool support.  In~2018, 3,603 benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and 5
benchmarks were excluded because of missing status information.

@@

Also shown in Figure~\ref{fig:job-pairs} is the total wall-clock time
(summed over all job pairs) for the competition.  This has more than
doubled, from 3.5~years in~2015 to 7.6~years in~2018.  (These are
solver run-times only.  Time that the cluster spent on pre- and
post-processing of benchmarks and solver output or on other auxiliary
tasks is not included.)  The increase in~2016 is largely due to the
inclusion of (non-incremental) benchmarks with unknown status; these
were evaluated with a wall-clock time limit of 10~minutes.  In~2017,
when these benchmarks had become eligible for the main track,
main-track job pairs were first run with a time limit of 10~minutes.
When it became apparent to the organizers that there were sufficient
resources to increase this limit, job pairs that had timed out were
re-run with a time limit of 20~minutes.  (In this way, about 2.9 years
of wall-clock time in 2017 were spent on job pairs that were then
re-run.  This time is not included in Figure~\ref{fig:job-pairs}.)
Table~\ref{table:time-limits} shows the wall-clock time limits that
were imposed on job pairs for each track and year.

\begin{table}
  \caption{Wall-clock time limits}
  \label{table:time-limits}
  \centering
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    & \multicolumn{1}{c|}{2015} & \multicolumn{1}{c|}{2016} & \multicolumn{1}{c|}{2017} & \multicolumn{1}{c|}{2018} \\
    \hline
    Main track        & 40~min & 40~min & 20~min\protect\footnotemark & 20~min \\
    Application track & 40~min & 40~min & 40~min\phantom{$^{1.}$} & 40~min \\
    Unsat-core track  &     -- & 40~min & 40~min\phantom{$^{1.}$} & 40~min \\
    \hline
  \end{tabular}
\end{table}
\footnotetext{These job pairs were first run with a time limit of
  10~minutes.  See the text for details.}

@@

Table~\ref{table:participants} lists the solvers that participated
between~2015 and~2018, together with their submitter(s) and
institutional affiliations (which may have changed over time).  Note
that most solvers were developed by a group of people; the table does
not list all contributors but only submitters.
%
Participants are also encouraged to submit a short system description
of their solver, including a list of its authors and an explanation of
the SMT solving approach used.  System descriptions are published on
the competition website.  They were submitted for about 60\,\% of all
solver versions since 2015.

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{11.7cm}|} %TODO: fix width
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC~\cite{DBLP:conf/cav/BraytonM10} & Valeriy Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE~\cite{DBLP:conf/cade/GieslBEFFOPSSST14} & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector~\cite{DBLP:conf/cav/NiemetzPWB18} & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI~\cite{colibri} & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo         & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3~\cite{cvc3}  & Kshitij Bansal
                        (New York University)\\
    CVC4~\cite{DBLP:conf/cav/BarrettCDHJKRT11}  & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP~\cite{maplestp} & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink~\cite{minkeyrink} & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2~\cite{opensmt2} & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB~\cite{prob}  & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B~\cite{q3b}    & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT~\cite{rasat} & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog~\cite{redlog} & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT~\cite{smt-rat} & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol~\cite{} & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT~\cite{} & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP~\cite{}        & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt~\cite{}     & Masahiro Sakai
                        (Toshiba Corporation)\\
    Vampire~\cite{}    & Giles Reger
                        (University of Manchester)\\
    veriT~\cite{DBLP:conf/cade/BoutonODF09} & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat~\cite{}       & Martin Velez
                        (University of California, Davis)\\
    Yices2~\cite{}     & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3~\cite{DBLP:conf/tacas/MouraB08} & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}
\end{table}

@@

The competition-wide ranking for the main track was retained and
refined after~2014.  Table~\ref{table:competition-wide-ranking} shows
the best three solvers according to this ranking for 2015--2018.  Also
shown is Z3, which ranked highly each year but was running hors
concours.  Sequential and parallel performance are reported
separately.
%
Note that scores are not directly comparable between different years:
the set of benchmarks changes from one year to the next, and also the
formula to compute the score changed (see below).  Moreover, readers
should keep in mind that this ranking---like any method that maps the
performance of solvers for a broad range of logics to a
one-dimensional value---is inherently limited.  We refer to the
individual division rankings (Section~\ref{sec:division-rankings}) for
more detailed information about each solver.

\begin{table}
  \caption{Competition-wide ranking for the main track: best solvers (by year)}
  \label{table:competition-wide-ranking}
  \centering
  \scalebox{.89}{\begin{tabular}{|c|p{2.2cm}|r|r|p{0cm}|c|p{2.2cm}|r|r|}
    \cline{1-4} \cline{6-9}
    \multicolumn{4}{|c|}{2015} & & \multicolumn{4}{c|}{2016} \\ \cline{1-4} \cline{6-9}
    Rank & Solver & Score  & Score  & & Rank & Solver & Score & Score \\
         &        & (sequential) & (parallel) & & & & (sequential) & (parallel) \\ \cline{1-4} \cline{6-9}
         & [Z3]              & 159.36 & 159.36 & &    & [Z3]   & 185.09 & 185.09 \\
      1. & CVC4              & 144.67 & 144.74 & & 1. & CVC4   & 180.95 & 181.19 \\
      2. & CVC4(exp)\footnotemark & 140.47 & 140.51 & & 2. & Yices2 & 119.29 & 119.29 \\
      3. & Yices             & 101.91 & 101.91 & & 3. & veriT  &  75.11 &  75.11 \\ \cline{1-4} \cline{6-9}
  \end{tabular}}

@@

Thirteen years after its inception, the SMT Competition is a
well-established event that garners wide interest from developers and
users of SMT solvers.  The competition continues to contribute to its
original goals, including adoption of the SMT-LIB format and sparking
further advances in SMT, and provides a yardstick of performance for
SMT solvers.

The last FLoC Olympiad was a period of significant growth for the SMT
Competition.  The average number of solver submissions increased
from~12 (for 2005--2014) to~26 (for 2015--2018).  The number of
SMT-LIB benchmarks used in the competition increased from~77,352
in~2014~\cite{CDW14} to 342,498 in 2018.  This is partly because new
benchmarks were added to the SMT Library, but also because benchmarks
with unknown status are now eligible for the competition, and because
the competition now uses all eligible benchmarks (rather than a
randomly selected subset).  Since the re-introduction of the
unsat-core track in~2016, the competition again has three separate
tracks.  This, together with new logics in the SMT Library, has
brought the total number of competition divisions to 115 in~2018, up
from just 40 divisions in~2014~\cite{CDW14}.  The number of job pairs
increased from 339,714 in~2014 to~1,776,062~in 2018, and the total
wall-clock time in~2018 exceeded 7.6~years.  By these measures, the
SMT Competition is one of the largest competitions in automated
reasoning.

@@

Despite the substantial amount of computational resources available to
the competition through StarExec, the competition has reached a size
where future organizers will have to carefully consider the number of
participants and benchmarks each year, to determine time limits that
will allow the computational workload to complete before the SMT
Workshop.

Further development of the competition should be undertaken with care,
keeping in mind that major changes require not just the competition
organizers, but also participants to adapt.  Immediate action items
for next year's organizers are
\begin{itemize}
\item the inclusion of {\tt check-sat} commands with unknown status in
  the application track, and
\item the definition of competition-wide rankings for the application
  and unsat-core track.
\end{itemize}

@@

Organizing the competition is a non-trivial amount of work.  For
instance, the competition tools (such as the benchmark scrambler and
trace executor) often require updating when new divisions are added,
or when changes are made to StarExec.  One could imagine adding
further tracks to the competition (e.g., for proof-producing solvers,
or for variants such as max-SMT), but the effort required to support
such tracks will need to be balanced against their added value.

To ensure that the competition results are practically relevant, it is
important that the SMT Library contains benchmarks that cover a wide
range of applications.  To this end, we again encourage users of SMT
solvers to submit new and interesting benchmarks to the library,
especially for underrepresented logics.  This is not just a valuable
contribution to the community, but also---since these benchmarks will
then be used in the competition---a straightforward way to incentivize
solver developers to tune their solvers accordingly.
